Missing Data: Current Practice in Football Research and Recommendations for

### 2020-11-17 paper/paper.Rmd
add date.
edit writing and formatting.
edits

\begin{center}
David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
\end{center}
\vspace{5mm}
**Affiliations**

@@

\vspace{5mm}
**Citation:** Borg DN, Nguyen R, and Tierney NJ. (2020). Missing Data: Current Practice in Football Research and Recommendations for Improvement.

@@

\vspace{5mm}
**Conflict of Interest**

The authors declare no conflicts of interest.

\vspace{5mm}
**Funding**

No funding was awarded for this project.

@@

*Unreported missing values.* While not always the case, even if
not reported, missing data can be obvious. For example, a study examined the
relationship between match performance indicators and outcome, in
Australian Football between the 2001 and 2016 [@Young2019], where at least one of the
91 performance indicators ('meters gained') included in the analysis was not available over the entire 2001-2016 period [@Young2019_mention_of_missing]. This was not noted anywhere in the paper, or in the supplementary
materials. It is unknown whether imputation was undertaken. Without
these details, it is difficult for other analysts and researchers to use
and/or extend the ideas presented in the paper, or reproduce the
analysis. It is also difficult for a reader to evaluate how the
missingness could bias the results. Authors should include a missing data statement,
irrespective of whether data are missing, or not.

@@

Much in the way of there is no single 'best' statistical method, there is no perfect, one-size-fits-all approach for imputing data. The goal is to generate similar values that might have been otherwise recorded. Sometimes this means using a neighborhood approach of finding similar values. Or it could mean predicting responses using linear or tree models. Other times the most likely value might have been 0 or the last or first value carried forward. For detailed descriptions, and a summary of these methods, we suggest [@Van_Buuren2012; @Schafer2002; @Cheema2014].

The imputation methods discussed so far impute a single value for each missing value, and have the eponymous name, "single imputation". Multiple imputation is a method where multiple values are imputed for each missing value, creating "m" datasets, which are then specially combined during analysis. Multiple imputation is generally seen as the best method to get the most reasonable inference from the data, as it reflects the uncertainty in the missing values. For more information on using multiple imputation, we recommend Stefan van Buuren's book [@Van_Buuren2012].

>>>>

\begin{center}
David N Borg(1,2), Robert Nguyen(3), Nicholas J Tierney(4,5)
\vspace{5mm}

17 November 2020
\vspace{5mm}

Preprint (not peer reviewed)
\end{center}
\vspace{5mm}
**Affiliations**

@@

\vspace{5mm}
**Citation:** Borg DN, Nguyen R, and Tierney NJ. (2020). Missing Data: Current Practice in Football Research and Recommendations for Improvement. *SportR$\chi$iv*. https://doi.org/xxxx

@@

\vspace{5mm}
**Conflict of Interest and Funding**

The authors declare no conflicts of interest. No funding was awarded for this project.

@@

*Unreported missing values.* While not always the case, even if
not reported, missing data can be obvious. For example, a study examined the
relationship between match performance indicators and outcome, in
Australian Football between the 2001 and 2016 [@Young2019], where at least one of the
91 performance indicators ('meters gained') included in the analysis was not available over the entire 2001-2016 period [@Young2019mentionofmissing]. This was not noted anywhere in the paper, or in the supplementary
materials. It is unknown whether imputation was undertaken. Without
these details, it is difficult for other analysts and researchers to use
and/or extend the ideas presented in the paper, or reproduce the
analysis. It is also difficult for a reader to evaluate how the
missingness could bias the results. Authors should include a missing data statement,
irrespective of whether data are missing, or not.

@@

Much in the way of there is no single 'best' statistical method, there is no perfect, one-size-fits-all approach for imputing data. The goal is to generate similar values that might have been otherwise recorded. Sometimes this means using a neighborhood approach of finding similar values. Or it could mean predicting responses using linear or tree models. Other times the most likely value might have been 0 or the last or first value carried forward. For detailed descriptions, and a summary of these methods, we suggest [@VanBuuren2012; @Schafer2002; @Cheema2014].

The imputation methods discussed so far impute a single value for each missing value, and have the eponymous name, "single imputation". Multiple imputation is a method where multiple values are imputed for each missing value, creating "m" datasets, which are then specially combined during analysis. Multiple imputation is generally seen as the best method to get the most reasonable inference from the data, as it reflects the uncertainty in the missing values. For more information on using multiple imputation, we recommend Stefan van Buuren's book [@VanBuuren2012].
